### 2021-10-02 docs/paper/sections/1_introduction.tex
Minor improvements to phrasing and layout

Open Source software (OSS) solutions have become more and more and important during the last decades.
Especially, this trend has been experiencing an additional updraft due to the ongoing spreading of OSS platforms such as GitHub or GitLab.
Open-source development offers many advantages as opposed to traditional closed-source development, including large numbers of volunteer contributions from the open-source community, increased transparency effects in security-related domains, and a high potential for reusing solutions.
These solutions are usually organized as \emph{packages} each of which attempts to solve an isolated problem and provide a generic interface.
Most commonly, packages are developed in a \emph{code repository} that is managed using a \emph{development platform} (such a \emph{GitHub}\footnote{\url{https://github.com/}}, \emph{GitLab}\footnote{\url{https://gitlab.com/}}, or \emph{Bitbucket}\footnote{\url{https://bitbucket.com/}}) and deployed to a \emph{package repository} using a \emph{package manager} (such as \emph{PyPI}\footnote{\url{https://pypi.org/}} using \emph{pip}\footnote{\url{https://pip.pypa.io/}} for the Python programming language, the \emph{npm registry}\footnote{\url{https://www.npmjs.com/}} for JavaScript/TypeScript, or the \emph{NuGet Gallery}\footnote{\url{https://www.nuget.org/}} for languages from the .NET ecosystem).
Other solutions or packages can then \emph{depend} on existing packages.
These dependency relations form a large directed graph that connects major parts of the software world for each popular programming language ecosystem.

Despite this connectedness by design, however, the development process of many packages is still characterized by an isolated approach:
While OSS developers commonly submit tickets to or contribute patches against \emph{upstream repositories} that they depend on to solve subproblems, the reverse direction of these edges -- called \emph{downstream dependencies} -- is often neglected by package developers when they extend, restructure, or refactor their solutions.
This knowledge deficit can cause a wide range of alignment issues, including poorly suited interfaces \citep{piccioni2013empirical}, unidentified defects \citep{wong2017more}, and compatibility problems \citep{bogart2015breaks}.
Eventually, all these issues impair the capabilities of the global OSS community to build and support high-quality products.
In the following, downstream dependencies will also be referred to shortly as \emph{dependencies}.

@@

The rest of this paper is organized as follows:
In \cref{sec:related_work}, we give a summary of existing approaches in this field.
In \cref{sec:framework}, we outline the overall conditions and the general approach for our solution to mining downstream dependencies.
In \cref{sec:dependency_collection} and \cref{sec:usage_mining}, we explain our approach for collecting downstream dependency repositories resp. mining usage data from these repositories in detail.
In \cref{sec:implementation}, we describe an implementation of the proposed approach and our design decisions for displaying usage samples to the user.
In \cref{sec:evaluation}, we examine the fitness of our data mining approach and the usability of the display solution.

>>>>

Open Source software (OSS) solutions have become more and more and important during the last decades.
Especially, this trend has been experiencing an additional updraft due to the ongoing spreading of OSS platforms such as GitHub or GitLab.
Open-source development offers many advantages as opposed to traditional closed-source development, including large numbers of volunteer contributions from the open-source community, increased transparency effects in security-related domains, and a high potential for reusing solutions.
These solutions are usually organized as \emph{packages} each of which attempts to solve an isolated problem and provide a generic interface.
Most commonly, packages are developed in a \emph{code repository} that is managed using a \emph{development platform} (such a \emph{GitHub}\footnote{\url{https://github.com/}}, \emph{GitLab}\footnote{\url{https://gitlab.com/}}, or \emph{Bitbucket}\footnote{\url{https://bitbucket.com/}}) and deployed to a \emph{package repository} using a \emph{package manager} (such as \emph{PyPI}\footnote{\url{https://pypi.org/}} using \emph{pip}\footnote{\url{https://pip.pypa.io/}} for the Python programming language, the \emph{npm registry}\footnote{\url{https://www.npmjs.com/}} for JavaScript/TypeScript, or the \emph{NuGet Gallery}\footnote{\url{https://www.nuget.org/}} for languages from the .NET ecosystem).
Other solutions or packages can then \emph{depend} on existing packages by declaring them in their manifest, i.e., the \code{package.json} file of an npm package, or the \code{requirements.txt} file of a Python project.
These dependency relations form a large directed graph that connects major parts of the software world for each popular programming language ecosystem.

Despite this connectedness by design, however, the development process of many packages is still characterized by an isolated approach:
While OSS developers commonly submit tickets to or contribute patches against \emph{upstream repositories} that they depend on to solve subproblems, the reverse direction of these edges -- called \emph{downstream dependencies} -- is often neglected by package developers when they extend, restructure, or refactor their solutions.
This knowledge deficit can cause a wide range of alignment issues, including poorly suited interfaces \citep{piccioni2013empirical}, unidentified defects \citep{wong2017more}, and compatibility problems \citep{bogart2015breaks}.
Eventually, all these issues impair the capabilities of the global OSS community to build and support high-quality products.
In the following, we also refer to downstream dependencies will shortly as \emph{dependencies}.

@@

The rest of this paper is organized as follows:
In \cref{sec:related_work}, we give a summary of existing approaches in this field.
In \cref{sec:framework}, we outline the overall conditions and the general approach for our solution to mining downstream dependencies.
In \cref{sec:dependency_collection} and \cref{sec:usage_mining}, we explain our approach for collecting downstream dependency repositories and mining usage data from these repositories in detail, resp.
In \cref{sec:implementation}, we describe an implementation of the proposed approach and our design decisions for displaying usage samples to the user.
In \cref{sec:evaluation}, we examine the fitness of our data mining approach and the usability of the display solution.

### 2021-10-02 docs/paper/sections/2_related_work.tex
Minor improvements to phrasing and layout

A common purpose for \emph{analyzing dependency graphs} is to discover transitive (indirect) upstream dependencies of a project and assess their impact on the stability and vulnerability of the project.
This is done by \citet{kikas2017structure} which apply their methodology to the JavaScript, Ruby, and Rust ecosystems and gain statistical insights into the centrality and evolution of these ecosystems.
\citet{decan2018impact} pursue a similar approach in order to measure the spreading of security vulnerabilities along the chain of downstream dependencies.

@@

Dependency graphs do not provide information about package usage on a finer granularity, so reference insights about single package members such as classes or methods remain unknown.
The process of extracting fine-granular information about all references to individual elements of an interface is often referred to as \emph{API usage analysis}.
A simple approach to this problem is to perform a string search for package names or members in the dependencies \citep{mileva2010mining}.
\citet{qiu2016understanding} propose a more elaborated approach that operates on the \emph{abstract syntax tree} (AST) of each parsed dependency to avoid false positive matches that can be caused by ambivalent identifier names.
A very similar approach is also chosen by \citet{sawant2017fine} who evaluate the collected usage data in order to analyze the historical importance of certain features supported by an API.

@@

\citet{collard2013srcml} propose an infrastructure called \caps{srcML} that is aimed to create a unified representation of multilingual source code snippets for arbitrary analysis purposes, including the construction of call graphs.
Another solution is proposed by \citet{bogar2018lightweight} who focus on analyzing multilingual codebases and who utilize an island parser to build call graphs.
Furthermore, \citet{antal2018static} compare several call graph generators for JavaScript.
They also emphasize that for languages supporting dynamic typing or meta-programming mechanisms, static call graph generators have a limited precision by design.

To bring together dependency graphs and call graphs, \emph{ecosystem call graphs} can be constructed by applying call graphs to entire ecosystems, crossing repository boundaries.
\citet{hejderup2018software} propose an approach to do so for the JavaScript/npm ecosystem in which they also model the semantic version numbers of packages and perform an impact analysis for vulnerabilities based on their graph; a very similar method is described by \citet{boldi2020fine} to formulate the vision of new design for package management systems that provides options for finer-granular reuse of solutions.
\citet{nielsen2021modular} elaborate on this work by developing a call graph generator tailored to this problem that comes with significant improvements in terms of recall and performance.
\citet{wang2020empirical} describe a similar approach for the same ecosystem and apply it to the domain of locating security issues in a project's upstream dependencies.
With \caps{Pr\"azi} (not yet peer-reviewed), \citet{hejderup2021praezi} also describe their implementation of a dependency-scale call graph for the Rust/Cratesio ecosystem in-depth.
For Java/Maven, \citet{keshani2021scalable} proposes another solution.

After collecting these usage data, additional processing is possible to extract general usage information from the extensive raw data.
Next to straightforward grouping and counting operations, such aggregations can be built by detecting popular \emph{usage patterns} of API features in the downstream dependencies.
\citet{zhong2009mapo} propose a framework to do so that, for instance, finds sequences of API members that are invoked frequently and even uses these patterns for guiding API users by giving them recommendations.
\citet{hanam2019aiding} pursue a different goal in their tool that helps package developers assess the impact of breaking API changes on the functionality of downstream dependencies.

@@

To establish an adequate developer experience, any collected usage data still need to be presented suitably.
A common approach for this is a hierarchical and navigatable representation of the collected call tree.
The first form of this representation has been invented as the interactive ``message set'' tool for browsing senders and implementors in an object-oriented system for the Smalltalk programming environment \citep[section~10.1f.]{goldberg1984smalltalk}.
Modern alternatives include the \caps{Stacksplorer} \citep{karrer2011stacksplorer} or \caps{Blaze} \citep{kramer2012blaze} that list next to the currently focused method other methods which are adjacent to this method in the call graph.
With a focus on exploring references to API members, \citet{de2013multi} propose a set of additional views, including hierarchical lists and word clouds for highly referenced identifiers.
Focusing on exploring APIs in a non-code-centric way, \citet{hora2015apiwave} propose a dashboard that allows developers to browse existing APIs and track possibly breaking changes in their interfaces.

>>>>

A common purpose for \emph{analyzing dependency graphs} is to discover transitive (indirect) upstream dependencies of a project and assess their impact on the stability and vulnerability of the project.
This is done by \citet{kikas2017structure} which apply their methodology to the JavaScript, Ruby, and Rust ecosystems and gain statistical insights into the centrality and evolution of these ecosystems.
\citet{decan2018impact} pursue a similar approach to measure the spreading of security vulnerabilities along the chain of downstream dependencies.

@@

Dependency graphs do not provide information about package usage on a finer granularity, so reference insights about single package members such as classes or methods remain unknown.
The process of extracting fine-grained information about all references to individual elements of an interface is often referred to as \emph{API usage analysis}.
A simple approach to this problem is to perform a string search for package names or members in the dependencies \citep{mileva2010mining}.
\citet{qiu2016understanding} propose a more elaborated approach that operates on the \emph{abstract syntax tree} (AST) of each parsed dependency to avoid false positive matches caused by ambivalent identifier names.
A very similar approach is also chosen by \citet{sawant2017fine} who evaluate the collected usage data to analyze the historical importance of certain features supported by an API.

@@

\citet{collard2013srcml} propose an infrastructure called \caps{srcML} that aims to create a unified representation of multilingual source code snippets for arbitrary analysis purposes, including the construction of call graphs.
Another solution is proposed by \citet{bogar2018lightweight} who focus on analyzing multilingual codebases and who utilize an island parser to build call graphs.
Furthermore, \citet{antal2018static} compare several call graph generators for JavaScript.
They also emphasize that for languages supporting dynamic typing or meta-programming mechanisms, static call graph generators have a limited precision by design.

To bring together dependency graphs and call graphs, \emph{ecosystem call graphs} can be constructed by applying call graphs to entire ecosystems, crossing repository boundaries.
\citet{hejderup2018software} propose an approach to do so for the JavaScript/npm ecosystem in which they also model the semantic version numbers of packages and conduct an impact analysis for vulnerabilities based on their graph; a very similar method is described by \citet{boldi2020fine} to formulate the vision of a new design for package management systems that provides options for finer-grained reuse of solutions.
\citet{nielsen2021modular} elaborate on this work by developing a call graph generator tailored to this problem that comes with significant improvements in terms of recall and performance.
\citet{wang2020empirical} describe a similar approach for the same ecosystem and apply it to the domain of locating security issues in a project's upstream dependencies.
With \caps{Pr\"azi} (not yet peer-reviewed), \citet{hejderup2021praezi} also describe their implementation of a dependency-scale call graph for the Rust/Cratesio ecosystem in-depth.
For Java/Maven, \citet{keshani2021scalable} proposes another solution.
Another approach is implemented in the \emph{precise code intelligence} feature of the Sourcegraph code search engine that makes code references explorable across repository boundaries for repositories that provide a metadata file in the \emph{Language Server Index Format}.\footnote{\url{https://docs.sourcegraph.com/code_intelligence/explanations/precise_code_intelligence}}

After collecting these usage data, additional processing is possible to extract general usage information from the extensive raw data.
Next to trivial grouping and counting operations, such aggregations can be built by detecting popular \emph{usage patterns} of API features in the downstream dependencies.
\citet{zhong2009mapo} propose a framework to do so that, for instance, finds sequences of API members that are invoked frequently and even uses these patterns for guiding API users by giving them recommendations.
\citet{hanam2019aiding} pursue a different goal in their tool that helps package developers assess the impact of breaking API changes on the functionality of downstream dependencies.

@@

To establish an adequate developer experience, any collected usage data still need to be presented suitably.
A common approach for this is a hierarchical and navigatable representation of the collected call tree.
An early form of this representation has been invented as the interactive ``message set'' tool for browsing senders and implementors in an object-oriented system for the Smalltalk programming environment \citep[section~10.1f.]{goldberg1984smalltalk}.
Modern alternatives include the \caps{Stacksplorer} \citep{karrer2011stacksplorer} or \caps{Blaze} \citep{kramer2012blaze} that list next to the currently focused method other methods which are adjacent to this method in the call graph.
With a focus on exploring references to API members, \citet{de2013multi} propose a set of additional views, including hierarchical lists and word clouds for highly referenced identifiers.
Focusing on exploring APIs in a non-code-centric way, \citet{hora2015apiwave} propose a dashboard that allows developers to browse existing APIs and track possibly breaking changes in their interfaces.

### 2021-10-02 docs/paper/sections/3_framework.tex
Minor improvements to phrasing and layout

Furthermore, we pose two organizational requirements to the tool solution to ensure its practical applicability for the target user group:

\begin{enumerate}[label=R\arabic*]
	\item \label[requirement]{req1} The tool runs out-of-the-box with a small footprint, without the user needing to perform any sophisticated setup steps or providing additional system resources for it to work.
	\item \label[requirement]{req2} The tool blends in with the usual workflow of package developers at the best possible rate.
\end{enumerate}

>>>>

Furthermore, we pose two organizational requirements to the eventual tool solution to ensure its practical applicability for the target user group:

\begin{enumerate}[label=R\arabic*]
	\item \label[requirement]{req1} The tool runs out-of-the-box with a small footprint, without the user needing to perform any sophisticated setup steps or providing additional system resources.
	\item \label[requirement]{req2} The tool blends in with the usual workflow of package developers at the best possible rate.
\end{enumerate}

### 2021-10-02 docs/paper/sections/4_dependency_collection.tex
Minor improvements to phrasing and layout

In the first step, a list of repositories that depend on the target package has to be assembled.
As mentioned in \cref{sec:related_work/usage_samples}, many approaches start with a large downloaded corpus of unspecific repositories from the ecosystem which they then iterate over to filter the relevant repositories.
However, a major drawback of this approach is the high resource demand for both creating and traversing this corpus.
This approach is suited for analyzing repositories in large scale but is not conform to \cref{req1} due to the large footprint in terms of computational power and time.

@@

\begin{enumerate}[label=(\roman*)]
	\item Package repositories that maintain a doubly-connected edge list of all packages in an ecosystem depending on each other.
	\item Code search engines that index the source code of many repositories from OSS platforms.
		Using these search engines, we can query all repositories that declare a dependency on the target package in their package metadata file.
\end{enumerate}

>>>>

In the first step, a list of repositories that depend on the target package has to be assembled.
As mentioned in \cref{sec:related_work/usage_samples}, many approaches start with a large downloaded corpus of unspecific repositories from the ecosystem, which they then iterate over to filter the relevant repositories.
However, a major drawback of this approach is the high resource demand for both creating and traversing this corpus.
This approach is suited for analyzing repositories on a large scale but is not conform to \cref{req1} because of its large footprint in terms of computational power and time.

@@

\begin{enumerate}[label=(\roman*)]
	\item Package repositories that maintain a doubly-connected edge list of all interdependent packages in an ecosystem.
	\item Code search engines that index the source code of many repositories from OSS platforms.
		Using these search engines, we can query all repositories that declare a dependency on the target package in their package manifest file.
\end{enumerate}

### 2021-10-02 docs/paper/sections/5_usage_mining.tex
Minor improvements to phrasing and layout

Having downloaded the selected downstream dependency repositories, we can proceed to extract usage samples for the target package from each dependency repository.
Our goal is to identify these usage samples on a fine-granular level so that we can trace them down to the single identifiers that were exposed by the target package.

@@

In a second step, a static type analysis is performed against each dependency ASF together with the target package's ASF.
The results of the type analysis are attached to the ASF so that every identifier is annotated with a type symbol that links to the declaration of the type for the identifier.
For instance, after this step, every variable node will contain a link to the assignment node of this variable, and every function call expression will contain a link to the definition of this function. \TODO{Give example?}
The operating principle of this type analysis depends on the kind of programming language; i.e., in statically typed languages, type symbols can usually be retrieved from the declaration of an identifier, whereas in dynamically typed languages, a control flow analysis will be required to identify the origin of every identifier's type.

In the final step, from each ASF all nodes are collected whose type is declared in the target package.
To identify these links, we define a set of language-specific patterns for AST subtrees that constitute a usage expression.
In particular, each of these patterns expects a node containing a link to an identifier declaration (see \cref{fig:usage_mining/patterns}).

@@

\begin{algorithm}
	\caption{Extraction of usage samples.}\label{alg:usage_mining}

>>>>

Having downloaded the selected downstream dependency repositories, we can proceed to extract usage samples for the target package from each dependency repository.
Our goal is to collect fine-grained source code excerpts that reference single identifiers exposed by the target package.

\begin{algorithm}[b]
	\caption{Extraction of usage samples.}\label{alg:usage_mining}

@@

In a second step, we perform a static type analysis against each dependency ASF together with the target package's ASF.
The results of the type analysis are attached to the ASF so that every identifier is annotated with a type symbol that links to the declaration of the type for the identifier.
For instance, after this step, every variable node will contain a link to the assignment node of this variable, and every function call expression will contain a link to the definition of this function.
The operating principle of this type analysis depends on the kind of programming language; i.e., in statically typed languages, type symbols can usually be retrieved from the declaration of an identifier, whereas in dynamically typed languages, a control flow analysis will be required to identify the origin of every identifier's type.

In the final step, from each ASF, we collect all nodes whose type is declared in the target package.
To identify these links, we define a set of language-specific patterns for AST subtrees that constitute a usage expression.
In particular, each of these patterns expects a node containing a link to an identifier declaration (see \cref{fig:usage_mining/patterns}).

@@

### 2021-10-02 docs/paper/sections/6_implementation.tex
Minor improvements to phrasing and layout

In our implementation, we are restraining ourselves to analyzing node.js projects written in JavaScript.
This decision is based on the large prevalence of JavaScript\footnote{GitHut 2.0: A small place to discover languages in GitHub. \url{https://madnight.github.io/githut/\#/pushes}} as well as the first-rate extent of the npm ecosystem\footnote{Modulecounts: \url{http://www.modulecounts.com/}} which has been unsurpassed since 2015.
In the following, we are describing how we apply the single steps of the approach presented above to this particular domain.

@@

The de facto standard package repository for node.js packages is npm.\footnote{\url{https://www.npmjs.com/}}
To fetch downstream dependency packages from the npm registry, we utilize the npm package npm-dependants\footnote{\url{https://www.npmjs.com/package/npm-dependants}} which scrapes the dependents list from npmjs.com as there does not exist a publicly available API for these data.
To keep the footprint of our implementation small, we only scrape a part of the dependents list.
We then request basic metadata for each found package from the npm registry, in particular including a URL to the source code archive of the latest package version (commonly referred to as \emph{tarball}), by using the package-metadata package.\footnote{\url{https://www.npmjs.com/package/package-metadata}}
Finally, we download and extract each source code archive from the npm registry with the help of the download-package-tarball package.\footnote{\url{https://www.npmjs.com/package/download-package-tarball}}

\begin{lstlisting}[language=json,
	label=sec:implementation/package.json,
	caption={An example \code{package.json} manifest file for the \code{ripple-lib} package specifying multiple dependencies (truncated).},
	float=tpb,
	floatplacement=tbp,
	abovecaptionskip=-5pt
]
{
	"name": "ripple-lib",
	"version": "1.10.0",
	"description": "A TypeScript/JavaScript API for interacting with the XRP Ledger",
	"dependencies": {
		"bignumber.js": "^9.0.0",
		"https-proxy-agent": "^5.0.0",
		"jsonschema": "1.2.2"
	}
}
\end{lstlisting}

For the second approach, using a code search engine, we have identified Sourcegraph\footnote{\url{https://sourcegraph.com/}} as a promising platform.
npm packages specify their dependencies in a package.json metadata file (see \cref{sec:implementation/package.json}).
To find all packages depending on a package \code{<pkg>}, we can search all available package.json files for the substring \code{"<pkg>":}.
Refined with some additional constraints to filter out package copies from \code{node\_modules} folders that downstream developers have pushed accidentally, and to limit the number of results retrieved, our final Sourcegraph query has the following form:

@@

We submit this search query by sending it to the publicly available Sourcegraph GraphQL API.\footnote{\url{https://docs.sourcegraph.com/api/graphql}}
For each hit, we request additional information from the API about the matched repository, including the name of the depending package and the URL to the original repository.
We use the npm package graphql-request\footnote{\url{https://www.npmjs.com/package/graphql-request}} for accessing the GraphQL API.
After retrieving a search result, we download a snapshot of the full repository by using the download-git-repo package\footnote{\url{https://www.npmjs.com/package/download-git-repo}} which combines access to the APIs of different code host providers such as GitHub or BitBucket.

@@

In our solution, we are focusing on mining usage samples from downstream dependencies that are written in JavaScript or TypeScript.
\TODO{Is this the right place to introduce TypeScript?}
TypeScript is an extension of JavaScript that adds strong typing to the former which is based on a combination of implicit type inference and explicit type declarations.
As the type-checking capabilities of the TypeScript compiler are optional and can be turned off via configuration, we consider the JavaScript syntax a subset of TypeScript syntax.

For parsing the dependencies, we utilize the public TypeScript Compiler API.\footnote{\url{https://github.com/Microsoft/TypeScript/wiki/Using-the-Compiler-API}}
The TypeScript compiler's \emph{binder} component annotates every node of the parsed program with a type symbol if it can resolve one.
After performing a type analysis of the dependency by instrumenting the binder, we collect all nodes from the type-annotated ASF that refer to the target package by traversing it with a visitor function.
This visitor function matches each node against a list of defined patterns (similar to \cref{fig:usage_mining/patterns}) based on the \code{syntaxKind}s of the node and its ancestors and descendants and then looks up the declaration of the node's type symbol to compare it with the source directory of the target package.

@@

@@

After all references have been collected, a proper UI is still required to provide users easy access to these data that fulfills the requirements mentioned above.
To satisfy \cref{req2} and make our tool available in the usual working environment of users, we implement it as an extension to a popular IDE.
One of the most popular IDEs that support JavaScript, and the one with the highest annual growth, is Visual Studio Code.\footnote{\url{https://code.visualstudio.com/}}\footnote{Top IDE Index by Pierre Carbonnelle (2021): \url{https://pypl.github.io/IDE.html}}
As it also provides a comprehensive set of APIs for extension developers, we have decided to roll our tool out as a VS Code extension.

\begin{figure*}[b]
	\centering
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\textwidth]{sections/6_implementation/extension/dependencies.png}
		\caption[LoF entry]{The dependency browser.}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\textwidth]{sections/6_implementation/extension/references.png}
		\caption[LoF entry]{The usage browser.}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\textwidth]{sections/6_implementation/extension/codelens.png}
		\caption[LoF entry]{The CodeLens integration for the class \code{GraphQLError}.}
	\end{subfigure}

@@

Both references and members are organized in a tree view reflecting the hierarchical structure of the original software repository.
In addition, we recognize the effort of browsing large lists of dependency data and encounter it by providing an ``I'm feeling lucky'' button for every view that redirects the user to a random dependency resp. reference to gain a faster, unbiased impression of usage samples.

We implement each of these views in the VS Code Extension API (VSCE)\footnote{\url{https://code.visualstudio.com/api}}, we use the Tree View API and the CodeLens API.
One challenge was to deliver fast results and not to block the user interface because even though the lightweight mining approaches we chose to satisfy \cref{req1} above, querying, downloading, and analyzing each dependency takes a few seconds (see \cref{sec:evaluation/TODO}).
To overcome this issue, we use pagination wherever possible and applied the observer pattern to push incremental UI updates.
As the VSCE API per se does not provide for multithreading or multiprocessing operations but suggests a Promise/A+-driven dataflow, we adopt this style for our data model using the JavaScript concepts asynchronous functions, \code{AsyncIterator}s, and \code{Promise.all} to avoid busy waiting inside the data collection process.

>>>>

In our implementation, we confine ourselves to analyzing node.js projects written in JavaScript.
This decision is based on the large prevalence of JavaScript\footnote{GitHut 2.0: A small place to discover languages in GitHub. \url{https://madnight.github.io/githut/\#/pushes}} as well as the first-rate extent of the npm ecosystem\footnote{Modulecounts: \url{http://www.modulecounts.com/}} which has been unsurpassed since 2015.
In the following, we describe how we apply the single steps of the approach presented above to this particular domain.

@@

The de facto standard package repository for node.js is npm.\footnote{\url{https://www.npmjs.com/}}
To fetch downstream dependency packages from the npm registry, we use the npm package npm-dependants\footnote{\url{https://www.npmjs.com/package/npm-dependants}} which scrapes the dependents list from npmjs.com as there does not exist a publicly available API for these data.
To keep the footprint of our implementation small, we only scrape a part of the dependents list.
We then request basic metadata for each found package from the npm registry, in particular including a URL to the source code archive of the latest package version (commonly referred to as \emph{tarball}), by using the package-metadata package.\footnote{\url{https://www.npmjs.com/package/package-metadata}}
Finally, we download and extract each source code archive from the npm registry with the help of the download-package-tarball package.\footnote{\url{https://www.npmjs.com/package/download-package-tarball}}

\begin{lstlisting}[language=json,
	label=sec:implementation/package.json,
	caption={An example \code{package.json} manifest file for the \code{ripple-lib} package specifying multiple dependencies (truncated).},
	float=tpb,
	floatplacement=tbp,
	abovecaptionskip=-5pt
]
{
	"name": "ripple-lib",
	"version": "1.10.0",
	"description": "An API for the XRP Ledger",
	"dependencies": {
		"bignumber.js": "^9.0.0",
		"https-proxy-agent": "^5.0.0",
		"jsonschema": "1.2.2"
	}
}
\end{lstlisting}

For the second approach, using a code search engine, we have identified Sourcegraph\footnote{\url{https://sourcegraph.com/}} as a promising platform.
npm packages specify their dependencies in a package.json metadata file (see \cref{sec:implementation/package.json}).
To find all packages depending on a package with the name \code{<pkg>}, we can search all available package.json files for the substring \code{"<pkg>":}.
Refined with some additional constraints to filter out package copies from \code{node\_modules} folders that downstream developers have pushed accidentally, and to limit the number of results retrieved, our final Sourcegraph query has the following form:

@@

We submit this search query by sending it to the publicly available Sourcegraph GraphQL API\footnote{\url{https://docs.sourcegraph.com/api/graphql}} by using the npm package graphql-request\footnote{\url{https://www.npmjs.com/package/graphql-request}}.
For each hit, we request additional information from the API about the matched repository, including the name of the depending package and the URL to the original repository.
After retrieving a search result, we download a snapshot of the full repository by using the download-git-repo package\footnote{\url{https://www.npmjs.com/package/download-git-repo}} which unifies access to the APIs of different code host providers such as GitHub or BitBucket.

@@

For parsing the dependencies, we utilize the public TypeScript Compiler API.\footnote{\url{https://github.com/Microsoft/TypeScript/wiki/Using-the-Compiler-API}}
TypeScript is an extension of JavaScript that adds strong typing to the former; typing is based on a combination of implicit type inference and explicit type declarations.
As the type-checking capabilities of the TypeScript compiler are optional and can be turned off via configuration, we consider the JavaScript syntax a subset of TypeScript syntax.
As a side benefit, our implementation is also able to analyze packages implemented in true TypeScript.

The TypeScript compiler's \emph{binder} component annotates every node of the parsed program with a type symbol if it can resolve one.
After performing a type analysis of the dependency by instrumenting the binder, we collect all nodes from the type-annotated ASF that refer to the target package by traversing it with a visitor function.
This visitor function matches each node against a list of defined patterns (similar to \cref{fig:usage_mining/patterns}) based on the syntax kinds of the node, its ancestors, and its descendants, and then looks up the declaration of the node's type symbol to compare it with the source directory of the target package.

@@

\begin{figure*}[b]
	\centering
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\linewidth]{sections/6_implementation/extension/dependencies.png}
		\caption[LoF entry]{The dependency browser.}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\linewidth]{sections/6_implementation/extension/references.png}
		\caption[LoF entry]{The usage browser.}
		\label{fig:implementation/presentation/screenshot/usage_browser}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\linewidth]{sections/6_implementation/extension/codelens.png}
		\caption[LoF entry]{The CodeLens integration for the class \code{GraphQLError}.}
	\end{subfigure}

@@

After all references have been collected, a proper UI is still required to provide users easy access to these data that fulfills the requirements mentioned above.
To satisfy \cref{req2} and make our tool available in the usual working environment of users, we implement it as an extension to a popular IDE.
One of the most popular IDEs that support JavaScript, and the one with the highest annual growth, is Visual Studio Code.\footnote{\url{https://code.visualstudio.com/}}\footnote{Top IDE Index by Pierre Carbonnelle (2021): \url{https://pypl.github.io/IDE.html}}
As it also provides a comprehensive set of APIs for extension developers, we have decided to roll our tool out as a VS Code extension.\footnote{To support other researchers in reusing our solution, we also provide a CLI for our implementation.}

@@

Both references and members are organized in a tree view reflecting the hierarchical structure of the original software repository.
In addition, we recognize the effort of browsing large lists of dependency data and encounter it by providing an ``I'm feeling lucky'' button for every view that redirects the user to a random dependency or reference, resp., to gain a faster, unbiased impression of usage samples.

We implement each of these views in the VS Code Extension API (VSCE)\footnote{\url{https://code.visualstudio.com/api}}, we use the Tree View API and the CodeLens API.
One challenge has been to deliver fast results and not to block the user interface because even inspite of the lightweight mining approaches we chose to satisfy \cref{req1} above, querying, downloading, and analyzing each dependency takes a few seconds (see \cref{sec:evaluation}).
To overcome this issue, we use pagination wherever possible and apply the observer pattern to push incremental UI updates.
As the VSCE API per se does not provide for multithreading or multiprocessing operations but suggests a Promise/A+-driven dataflow, we adopt this style for our object model by using the JavaScript concepts asynchronous functions, \code{AsyncIterator}s, and \code{Promise.all} to avoid busy waiting inside the data collection process.
